A tool set for random number generation on GPUs in R

### 2022-01-02 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
p1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3085 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Equivalently, \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
\begin{CodeChunk}
\begin{CodeInput}
R> myStreamsGpu2 = createStreamsGpu(4)
\end{CodeInput} 
\end{CodeChunk}

@@

Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart. Streams created on GPU does not remain in the memory when a new \proglang{R} session begins. However, we can save the streams' current states and initial states in matrix form on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, the streams will restart from the current states after they are reloaded. In this way we can reproduce the exact same sequence of random numbers in simulations. Below is the code that saves streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
\begin{CodeChunk}
\begin{CodeInput}
R> saveRDS(as.matrix(createStreamsGpu(n = 4)), "myStreams.rds")
R> # Load the streams object as streams_saved
R> streams_saved <- vclMatrix(readRDS("myStreams.rds"))
\end{CodeInput} 
\end{CodeChunk}

@@

We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{clrng::rnorm}. The best performance we have seen using \fct{clrng::rnorm} is above 160 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size goes larger.
\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(n = 512 * 128)
R> system.time(clrng::rnorm(c(10000, 10000), streams = streams,
+    Nglobal = c(512, 128), type = "double"))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.050   0.000   0.051
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(n = 256 * 64)
R> month_gpu <- vclMatrix(month, type = "integer")
R> system.time(result_month <- clrng::fisher.sim(month_gpu, 1000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.311   0.006   0.317
\end{CodeOutput}
\begin{CodeInput}
R> result_month$threshold
\end{CodeInput}
\begin{CodeOutput}
## [1] -47955
\end{CodeOutput}
\begin{CodeInput}
R> result_month$simNum
\end{CodeInput}
\begin{CodeOutput}
## [1] 1015808
\end{CodeOutput}
\begin{CodeInput}
R> result_month$counts
\end{CodeInput}
\begin{CodeOutput}
## [1] 409429
\end{CodeOutput}
\begin{CodeInput}
R> result_month$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.403
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
R> system.time(result_week <- clrng::fisher.sim(week_GPU, 10000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   1.940   0.013   1.951
\end{CodeOutput}
\begin{CodeInput}
R> result_week$threshold
\end{CodeInput}
\begin{CodeOutput}
## [1] -54990
\end{CodeOutput}
\begin{CodeInput}
R> result_week$simNum
\end{CodeInput}
\begin{CodeOutput}
## [1] 10010624
\end{CodeOutput}
\begin{CodeInput}
R> result_week$counts
\end{CodeInput}
\begin{CodeOutput}
## [1] 1350
\end{CodeOutput}
\begin{CodeInput}
R> result_week$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.000135
\end{CodeOutput} 
\end{CodeChunk}

>>>>

Equivalently, \fct{createStreamsGpu} creates streams directly on the GPU and returns a `\code{vclMatrix}', which makes it slightly more efficient when the number of streams is large.
\begin{CodeChunk}
\begin{CodeInput}
R> rm(myStreamsCpu)
R> myStreamsGpu2 = createStreamsGpu(4)
\end{CodeInput} 
\end{CodeChunk}

@@

Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart. Streams created on GPU does not remain in the memory when a new \proglang{R} session begins. However, we can save the streams' current states and initial states in matrix form on the CPU in a file using the \fct{saveRDS} function, and later recall the streams with the \fct{readRDS} function, the streams will start from the current states after they are loaded. In this way we can reproduce the exact same sequence of random numbers in simulations. Below is the code that saves streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
\begin{CodeChunk}
\begin{CodeInput}
R> saveRDS(as.matrix(myStreamsGpu2), "myStreams.rds")
R> # Load the streams object as streams_saved
R> streams_saved <- vclMatrix(readRDS("myStreams.rds"))
\end{CodeInput} 
\end{CodeChunk}

@@

We generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{clrng::rnorm}. The best performance we have seen using \fct{clrng::rnorm} is above 160 times faster with $512 \times 128$ work-items than using the \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size goes larger.
\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(n = 512 * 128)
R> system.time(clrng::rnorm(c(10000, 10000), streams = streams,
+    Nglobal = c(512, 128), type = "double"))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.044   0.008   0.053
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(n = 256 * 64)
R> month_gpu <- vclMatrix(month, type = "integer")
R> system.time(result_month <- clrng::fisher.sim(month_gpu, 1000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.302   0.016   0.318
\end{CodeOutput}
\begin{CodeInput}
R> result_month$threshold
\end{CodeInput}
\begin{CodeOutput}
## [1] -47955
\end{CodeOutput}
\begin{CodeInput}
R> result_month$simNum
\end{CodeInput}
\begin{CodeOutput}
## [1] 1015808
\end{CodeOutput}
\begin{CodeInput}
R> result_month$counts
\end{CodeInput}
\begin{CodeOutput}
## [1] 409429
\end{CodeOutput}
\begin{CodeInput}
R> result_month$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.403
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
R> system.time(result_week <- clrng::fisher.sim(week_GPU, 10000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   1.937   0.017   1.952
\end{CodeOutput}
\begin{CodeInput}
R> result_week$threshold
\end{CodeInput}
\begin{CodeOutput}
## [1] -54990
\end{CodeOutput}
\begin{CodeInput}
R> result_week$simNum
\end{CodeInput}
\begin{CodeOutput}
## [1] 10010624
\end{CodeOutput}
\begin{CodeInput}
R> result_week$counts
\end{CodeInput}
\begin{CodeOutput}
## [1] 1350
\end{CodeOutput}
\begin{CodeInput}
R> result_week$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.000135
\end{CodeOutput} 
\end{CodeChunk}
